Quotients of the magmatic operad:\\ lattice structures and convergent rewrite systems

### 2018-09-07 Operads_rewrite.tex
Passe partie 1

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Nonsymmetric operads}
A \Def{nonsymmetric operad in the category of sets} (or a
\Def{nonsymmetric operad} for short) is a graded set
\begin{equation}
    \Oca = \bigsqcup_{n \geq 1} \Oca(n)
\end{equation}
together with maps
\begin{equation}
    \circ_i : \Oca(n) \times \Oca(m) \to \Oca(n + m - 1),
    \qquad 1 \leq i \leq n, 1 \leq m,
\end{equation}
called \Def{partial compositions}, and a distinguished element
$\Unit \in \Oca(1)$, the \Def{unit} of $\Oca$. This data has to satisfy,
for any $x \in \Oca(n)$, $y \in \Oca(m)$, and $z \in \Oca$, the three
relations
\begin{subequations}
\begin{equation} \label{equ:operad_axiom_1}
    (x \circ_i y) \circ_{i + j - 1} z = x \circ_i (y \circ_j z),
    \qquad
    1 \leq i \leq n, 1 \leq j \leq m,
\end{equation}
\begin{equation} \label{equ:operad_axiom_2}
    (x \circ_i y) \circ_{j + m - 1} z = (x \circ_j z) \circ_i y,
    \qquad
    1 \leq i < j \leq n,
\end{equation}
\begin{equation} \label{equ:operad_axiom_3}
    \Unit \circ_1 x = x = x \circ_i \Unit,
    \qquad 1 \leq i \leq n.
\end{equation}
\end{subequations}
Since we consider in this work only nonsymmetric operads, we shall call
these simply \Def{operads}.
\medbreak

Let us provide some elementary definitions and notations about operads.
If $x$ is an element of $\Oca$ such that $x \in \Oca(n)$ for an
$n \geq 1$, the \Def{arity} $|x|$ of $x$ is $n$. The
\Def{complete composition maps} of $\Oca$ are the map
\begin{equation}
    \circ : \Oca(n) \times
    \Oca\left(m_1\right) \times \dots \times \Oca\left(m_n\right)
    \to \Oca\left(m_1 + \dots + m_n\right)
\end{equation}
defined, for any $x \in \Oca(n)$ and $y_1, \dots, y_n \in \Oca$, by
\begin{equation} \label{equ:complete_composition}
    x \circ \left[y_1, \dots, y_n\right] :=
    \left(\cdots \left(\left(x \circ_n y_n\right)
    \circ_{n - 1} y_{n - 1}\right) \cdots\right) \circ_1 y_1.
\end{equation}
If $\Oca_1$ and $\Oca_2$ are two operads, a map
$\phi : \Oca_1 \to \Oca_2$ is an \Def{operad morphism} if it respects
the arities, sends the unit of $\Oca_1$ to the unit of $\Oca_2$, and
commutes with partial composition maps. A map $\phi : \Oca_1 \to \Oca_2$
is an \Def{operad antimorphism} if it respects the arities, sends the
unit of $\Oca_1$ to the unit of $\Oca_2$, and
\begin{math}
    \phi\left(x \circ_i y\right) = \phi(x) \circ_{|x|-i+1} \phi(y)
\end{math}
for all $x, y \in \Oca_1$ and $i \in [|x|]$. We say that $\Oca_2$ is a
\Def{suboperad} of $\Oca_1$ if $\Oca_2$ is a subset of $\Oca_1$
containing the unit of $\Oca_1$ and the partial composition maps of
$\Oca_2$ are the ones of $\Oca_1$ restricted on $\Oca_2$. For any subset
$\GeneratingSet$ of $\Oca$, the \Def{operad generated} by
$\GeneratingSet$ is the smallest suboperad $\Oca^\GeneratingSet$ of
$\Oca$ containing $\GeneratingSet$. When $\Oca = \Oca^\GeneratingSet$,
we say that $\GeneratingSet$ is a \Def{generating set} of $\Oca$. In
this case, and when additionally $\GeneratingSet$ is minimal among all
the subsets of $\Oca$ satisfying this property, $\GeneratingSet$ is a
\Def{minimal} generating set of $\Oca$. An \Def{operad congruence} is an
equivalence relation $\Congr$ on $\Oca$ such that $\Congr$ respects the
arities, and for any $x, x', y, y' \in \Oca$ such that $x \Congr x'$ and
$y \Congr y'$, $x \circ_i y$ is $\Congr$-equivalent to $x' \circ_i y'$
for any valid integer $i$. Given an operad congruence $\Congr$, the
\Def{quotient operad} $\Oca/_{\Congr}$ of $\Oca$ by $\Congr$ is the
operad of all $\Congr$-equivalence classes endowed with the partial
composition maps defined in the obvious way. In the case where all the
sets $\Oca(n)$, $n \geq 1$, are finite, the \Def{Hilbert series}
$\HilbertSeries_\Oca(t)$ of $\Oca$ is the series defined by
\begin{equation} \label{equ:hilbert_series}
    \HilbertSeries_\Oca(t) := \sum_{n \geq 1} \# \Oca(n) \; t^n.
\end{equation}
\medbreak

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Binary trees and magmatic operad}
A \Def{binary tree} is either the \Def{leaf} $\Leaf$ or a pair
$(\Tfr_1, \Tfr_2)$ of binary trees. We use the standard terminology
about binary trees (such as \Def{root}, \Def{internal node},
\Def{left child}, \Def{right child}, {\em etc.}) in this work. Let us
recall the main notions. The \Def{arity} $|\Tfr|$ (resp. \Def{degree}
$\Deg(\Tfr)$) of a binary tree $\Tfr$ is its number of leaves (resp.
internal nodes). A binary tree $\Tfr$ is \Def{quadratic} (resp.
\Def{cubic}) if $\Deg(\Tfr) = 2$ (resp. $\Deg(\Tfr) = 3$). We shall draw
binary trees the root to the top. For instance,
\begin{equation}
    \begin{tikzpicture}[xscale=.15,yscale=.18,Centering]
        \node(0)at(0.00,-4.50){};
        \node(2)at(2.00,-6.75){};
        \node(4)at(4.00,-6.75){};
        \node(6)at(6.00,-4.50){};
        \node(8)at(8.00,-4.50){};
        \node[NodeST](1)at(1.00,-2.25){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-4.50){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,-2.25){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(5);
        \draw[Edge](2)--(3);
        \draw[Edge](3)--(1);
        \draw[Edge](4)--(3);
        \draw[Edge](6)--(7);
        \draw[Edge](7)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(5.00,2){};
        \draw[Edge](r)--(5);
    \end{tikzpicture}
\end{equation}
is the graphical representation of the binary tree
\begin{math}
    ((\Leaf, (\Leaf, \Leaf)), (\Leaf, \Leaf)).
\end{math}
\medbreak

@@

    \begin{tikzpicture}[xscale=.15,yscale=.18,Centering]

@@

        \node(r)at(5.00,2){};

@@

    \begin{tikzpicture}[xscale=.15,yscale=.14,Centering]

@@

The operad $\Mag$ can be seen as the free operad generated by one binary
element~$\Product$. It satisfies the following universality property.
Let
\begin{math}
    \GeneratingSet := \GeneratingSet(2) := \left\{\Product'\right\}
\end{math}
be the graded set containing exactly one element $\Product'$ of arity
$2$. For any operad $\Oca$ and any map
\begin{math}
    f : \GeneratingSet(2) \to \Oca(2),
\end{math}
there exists a unique operad morphism $\phi : \Mag \to \Oca$ such that
$f = \phi \circ \Corolla$, where $\Corolla$ is the map sending
$\Product'$ to the unique binary tree of degree $1$ (and then, arity
$2$). In other terms, the diagram
\begin{equation}
    \begin{tikzpicture}[xscale=1.3,yscale=1.1,Centering]
        \node(G)at(0,0){\begin{math}\GeneratingSet\end{math}};
        \node(O)at(2,0){\begin{math}\Oca\end{math}};
        \node(AG)at(0,-2){\begin{math}\Mag\end{math}};
        \draw[Map](G)--(O)node[midway,above]{\begin{math}f\end{math}};
        \draw[Injection](G)--(AG)node[midway,left]
            {\begin{math}\Corolla\end{math}};
        \draw[Map,dashed](AG)--(O)node[midway,right]
            {\begin{math}\phi\end{math}};
    \end{tikzpicture}
\end{equation}
commutes.
\medbreak

@@

    \begin{tikzpicture}[xscale=.22,yscale=.22,Centering]

@@

    \begin{tikzpicture}[xscale=.19,yscale=.14,Centering]

@@

Let $\Rew$ be a set of rewrite rules and $\RewContext$ be the rewrite
relation induced by $\Rew$. Since $\RewContext$ is in particular a
binary relation on $\Mag$, the classical notations about closures apply
here: we denote by $\RewContextT$ (resp. $\RewContextRT$,
$\RewContextRST$) the transitive (resp. reflexive and transitive,
reflexive, symmetric, and transitive) closure of $\RewContext$.
\medbreak

When $\Tfr_0$, $\Tfr_1$, \dots, $\Tfr_k$ are binary trees such that
\begin{equation}
    \Tfr_0 \RewContext \Tfr_1 \RewContext \cdots \RewContext \Tfr_k,
\end{equation}
we say that $\Tfr_0$ is \Def{rewritable} by $\RewContext$ into $\Tfr_k$
in $k$ \Def{steps}. When there is no infinite chain
\begin{equation} \label{equ:infinite_chain}
    \Tfr_0 \RewContext \Tfr_1 \RewContext \Tfr_2 \RewContext \cdots
\end{equation}
we say that $\RewContext$ is terminating. To establish the termination
of a rewrite relation, we will use the following criterion.
\medbreak

\begin{Lemma}\label{lem:prefix_word_termination}
    Let $\Rew$ be a set of rewrite rules on $\Mag$. If for any
    $\Tfr, \Tfr' \in \Mag$ such that $\Tfr \Rew \Tfr'$ one has
    $\Tfr > \Tfr'$, then the rewrite relation induced by $\Rew$ is
    terminating.
\end{Lemma}
\begin{proof}
    Observe first that for any binary trees $\Tfr$ and $\Sfr$, the
    prefix word of $\Tfr \circ_i \Sfr$ is obtained by replacing the
    $i$th $\Zero$ of $\PrefixWord(\Tfr)$ by $\PrefixWord(\Sfr)$. For
    this reason, and due to the
    definition~\eqref{equ:complete_composition} of $\circ$, for any
    binary trees $\Sfr$ and $\Rfr_1, \dots, \Rfr_n$ where $n$ is the
    arity of $\Sfr$, the prefix word of
    $\Sfr \circ \left[\Rfr_1, \dots, \Rfr_n\right]$ is obtained by
    replacing from right to left each $\Zero$ of $\PrefixWord(\Sfr)$ by
    the prefix words of each $\Rfr_i$. This, together with the
    definition~\eqref{equ:rewrite_relation_induced} of the rewrite
    relation $\RewContext$ induced by $\Rew$ and the hypothesis of the
    statement of the lemma, implies that if $\Tfr$ and $\Tfr'$ are two
    binary trees such that $\Tfr \RewContext \Tfr'$,
    $\PrefixWord(\Tfr) > \PrefixWord\left(\Tfr'\right)$. This leads to
    the fact that any chain
    \begin{math}
        \Tfr_0 \Rew \Tfr_1 \Rew \Tfr_2 \Rew \cdots
    \end{math}
    is finite since
    \begin{math}
        \Tfr_0 > \Tfr_1 > \Tfr_2 > \cdots
    \end{math}
    and there is a finite number of binary trees of a fixed arity.
    Therefore, $\RewContext$ is terminating.
\end{proof}
\medbreak

@@

Let us now go back on operads. Let $\Congr$ be an operad congruence of
$\Mag$. If $\Tfr$ is a binary tree, we denote by $[\Tfr]_{\Congr}$ the
$\Congr$-equivalence class of $\Tfr$. By definition, $[\Tfr]_{\Congr}$
is an element of the quotient operad
\begin{equation} \label{equ:quotient_operad_mag}
    \Oca := \Mag/_{\Congr}.
\end{equation}
A set of rewrite rules $\Rew$ on $\Mag$ is an \Def{orientation} of
$\Congr$ if $\RewContextRST$ and $\Congr$ are equal as binary relations,
where $\RewContext$ is the rewrite relation induced by $\Rew$. Moreover,
$\Rew$ is a \Def{convergent} (resp. \Def{terminating}, \Def{confluent})
orientation of $\Congr$ if $\RewContext$ is convergent (resp.
terminating, confluent). We call \Def{presentation} of a quotient
operad $\Oca$ of the form~\eqref{equ:quotient_operad_mag} the data of
a generating set for the operad congruence $\Congr$. Observe that
any orientation of $\Congr$ is a presentation of $\Oca$, so that the
above nomenclature (\Def{convergent}, \Def{terminating}, and
\Def{confluent}) still holds for presentations. A presentation is said
to be \Def{finite} if it is a finite set.
\medbreak

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Nonsymmetric operads}
A \Def{nonsymmetric operad in the category of sets} (or a
\Def{nonsymmetric operad} for short) is a graded set
\begin{math}
    \Oca = \bigsqcup_{n \geq 1} \Oca(n)
\end{math}
together with maps
\begin{equation}
    \circ_i : \Oca(n) \times \Oca(m) \to \Oca(n + m - 1),
    \qquad 1 \leq i \leq n, 1 \leq m,
\end{equation}
called \Def{partial compositions}, and a distinguished element
$\Unit \in \Oca(1)$, the \Def{unit} of $\Oca$. This data has to satisfy,
for any $x \in \Oca(n)$, $y \in \Oca(m)$, and $z \in \Oca$, the three
relations
\begin{subequations}
\begin{equation} \label{equ:operad_axiom_1}
    (x \circ_i y) \circ_{i + j - 1} z = x \circ_i (y \circ_j z),
    \qquad
    1 \leq i \leq n, 1 \leq j \leq m,
\end{equation}
\begin{equation} \label{equ:operad_axiom_2}
    (x \circ_i y) \circ_{j + m - 1} z = (x \circ_j z) \circ_i y,
    \qquad
    1 \leq i < j \leq n,
\end{equation}
\begin{equation} \label{equ:operad_axiom_3}
    \Unit \circ_1 x = x = x \circ_i \Unit,
    \qquad 1 \leq i \leq n.
\end{equation}
\end{subequations}
Since we consider in this work only nonsymmetric operads, we shall call
these simply \Def{operads}.
\medbreak

Let us provide some elementary definitions and notations about operads.
If $x$ is an element of $\Oca$ such that $x \in \Oca(n)$ for an
$n \geq 1$, the \Def{arity} $|x|$ of $x$ is $n$. The
\Def{complete composition maps} of $\Oca$ are the map
\begin{equation}
    \circ : \Oca(n) \times
    \Oca\left(m_1\right) \times \dots \times \Oca\left(m_n\right)
    \to \Oca\left(m_1 + \dots + m_n\right)
\end{equation}
defined, for any $x \in \Oca(n)$ and $y_1, \dots, y_n \in \Oca$, by
\begin{equation} \label{equ:complete_composition}
    x \circ \left[y_1, \dots, y_n\right] :=
    \left(\cdots \left(\left(x \circ_n y_n\right)
    \circ_{n - 1} y_{n - 1}\right) \cdots\right) \circ_1 y_1.
\end{equation}
If $\Oca_1$ and $\Oca_2$ are two operads, a map
$\phi : \Oca_1 \to \Oca_2$ is an \Def{operad morphism} if it respects
the arities, sends the unit of $\Oca_1$ to the unit of $\Oca_2$, and
commutes with partial composition maps. A map $\phi : \Oca_1 \to \Oca_2$
is an \Def{operad antimorphism} if it respects the arities, sends the
unit of $\Oca_1$ to the unit of $\Oca_2$, and
\begin{math}
    \phi\left(x \circ_i y\right) = \phi(x) \circ_{|x|-i+1} \phi(y)
\end{math}
for all $x, y \in \Oca_1$ and $i \in [|x|]$. We say that $\Oca_2$ is a
\Def{suboperad} of $\Oca_1$ if $\Oca_2$ is a subset of $\Oca_1$
containing the unit of $\Oca_1$ and the partial composition maps of
$\Oca_2$ are the ones of $\Oca_1$ restricted on $\Oca_2$. For any subset
$\GeneratingSet$ of $\Oca$, the \Def{operad generated} by
$\GeneratingSet$ is the smallest suboperad $\Oca^\GeneratingSet$ of
$\Oca$ containing $\GeneratingSet$. When $\Oca = \Oca^\GeneratingSet$,
we say that $\GeneratingSet$ is a \Def{generating set} of $\Oca$. An
\Def{operad congruence} is an equivalence relation $\Congr$ on $\Oca$
such that $\Congr$ respects the arities, and for any
$x, x', y, y' \in \Oca$ such that $x \Congr x'$ and $y \Congr y'$,
$x \circ_i y$ is $\Congr$-equivalent to $x' \circ_i y'$ for any valid
integer $i$. Given an operad congruence $\Congr$, the
\Def{quotient operad} $\Oca/_{\Congr}$ of $\Oca$ by $\Congr$ is the
operad of all $\Congr$-equivalence classes endowed with the partial
composition maps defined in the obvious way. In the case where all the
sets $\Oca(n)$, $n \geq 1$, are finite, the \Def{Hilbert series}
$\HilbertSeries_\Oca(t)$ of $\Oca$ is the series defined by
\begin{equation} \label{equ:hilbert_series}
    \HilbertSeries_\Oca(t) := \sum_{n \geq 1} \# \Oca(n) \; t^n.
\end{equation}
\medbreak

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Binary trees and the magmatic operad}
A \Def{binary tree} is either the \Def{leaf} $\Leaf$ or a pair
$(\Tfr_1, \Tfr_2)$ of binary trees. We use the standard terminology
about binary trees (such as \Def{root}, \Def{internal node},
\Def{left child}, \Def{right child}, {\em etc.}) in this work. Let us
recall the main notions. The \Def{arity} $|\Tfr|$ (resp. \Def{degree}
$\Deg(\Tfr)$) of a binary tree $\Tfr$ is its number of leaves (resp.
internal nodes). A binary tree $\Tfr$ is \Def{quadratic} (resp.
\Def{cubic}) if $\Deg(\Tfr) = 2$ (resp. $\Deg(\Tfr) = 3$). We shall draw
binary trees the root to the top. For instance,
\begin{equation}
    \begin{tikzpicture}[xscale=.15,yscale=.16,Centering]
        \node(0)at(0.00,-4.50){};
        \node(2)at(2.00,-6.75){};
        \node(4)at(4.00,-6.75){};
        \node(6)at(6.00,-4.50){};
        \node(8)at(8.00,-4.50){};
        \node[NodeST](1)at(1.00,-2.25){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-4.50){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,-2.25){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(5);
        \draw[Edge](2)--(3);
        \draw[Edge](3)--(1);
        \draw[Edge](4)--(3);
        \draw[Edge](6)--(7);
        \draw[Edge](7)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(5.00,2.25){};
        \draw[Edge](r)--(5);
    \end{tikzpicture}
\end{equation}
is the graphical representation of the binary tree
\begin{math}
    ((\Leaf, (\Leaf, \Leaf)), (\Leaf, \Leaf)).
\end{math}
\medbreak

@@

    \begin{tikzpicture}[xscale=.15,yscale=.16,Centering]

@@

@@

    \begin{tikzpicture}[xscale=.15,yscale=.12,Centering]

@@

The operad $\Mag$ can be seen as the free operad generated by one binary
element~$\Product$. It satisfies the following universality property.
Let
\begin{math}
    \GeneratingSet := \GeneratingSet(2) := \left\{\Product'\right\}
\end{math}
be the graded set containing exactly one element $\Product'$ of arity
$2$. For any operad $\Oca$ and any map
\begin{math}
    f : \GeneratingSet(2) \to \Oca(2),
\end{math}
there exists a unique operad morphism $\phi : \Mag \to \Oca$ such that
$f = \phi \circ \Corolla$, where $\Corolla$ is the map sending
$\Product'$ to the unique binary tree of degree $1$ (and then, arity
$2$). In other terms, the diagram
\begin{equation}
    \begin{tikzpicture}[xscale=1.3,yscale=0.9,Centering]
        \node(G)at(0,0){\begin{math}\GeneratingSet\end{math}};
        \node(O)at(2,0){\begin{math}\Oca\end{math}};
        \node(AG)at(0,-2){\begin{math}\Mag\end{math}};
        \draw[Map](G)--(O)node[midway,above]{\begin{math}f\end{math}};
        \draw[Injection](G)--(AG)node[midway,left]
            {\begin{math}\Corolla\end{math}};
        \draw[Map,dashed](AG)--(O)node[midway,right]
            {\begin{math}\phi\end{math}};
    \end{tikzpicture}
\end{equation}
commutes.
\medbreak

@@

    \begin{tikzpicture}[xscale=.22,yscale=.21,Centering]

@@

    \begin{tikzpicture}[xscale=.22,yscale=.12,Centering]

@@

Let $\Rew$ be a set of rewrite rules and $\RewContext$ be the rewrite
relation induced by $\Rew$. Since~$\RewContext$ is in particular a
binary relation on $\Mag$, the classical notations about closures apply
here: we denote by $\RewContextT$ (resp. $\RewContextRT$,
$\RewContextRST$) the transitive (resp. reflexive and transitive,
and reflexive, symmetric, and transitive) closure of~$\RewContext$.
\medbreak

When $\Tfr_0$, $\Tfr_1$, \dots, $\Tfr_k$ are binary trees such that
\begin{equation}
    \Tfr_0 \RewContext \Tfr_1 \RewContext \cdots \RewContext \Tfr_k,
\end{equation}
we say that $\Tfr_0$ is \Def{rewritable} by $\RewContext$ into $\Tfr_k$
in $k$ \Def{steps}. When there is no infinite chain
\begin{equation} \label{equ:infinite_chain}
    \Tfr_0 \RewContext \Tfr_1 \RewContext \Tfr_2 \RewContext \cdots
\end{equation}
we say that $\RewContext$ is \Def{terminating}. To establish the
termination of a rewrite relation, we will use the following criterion.
\medbreak

\begin{Lemma}\label{lem:prefix_word_termination}
    Let $\Rew$ be a set of rewrite rules on $\Mag$. If for any
    $\Tfr, \Tfr' \in \Mag$ such that $\Tfr \Rew \Tfr'$ one has
    $\Tfr > \Tfr'$, then the rewrite relation induced by $\Rew$ is
    terminating.
\end{Lemma}
\begin{proof}
    Observe first that for any binary trees $\Tfr$ and $\Sfr$, the
    prefix word of $\Tfr \circ_i \Sfr$ is obtained by replacing the
    $i$th $\Zero$ of $\PrefixWord(\Tfr)$ by $\PrefixWord(\Sfr)$. For
    this reason, and due to the
    definition~\eqref{equ:complete_composition} of $\circ$, for any
    binary trees $\Sfr$ and $\Rfr_1, \dots, \Rfr_n$ where $n$ is the
    arity of $\Sfr$, the prefix word of
    $\Sfr \circ \left[\Rfr_1, \dots, \Rfr_n\right]$ is obtained by
    replacing from right to left each $\Zero$ of $\PrefixWord(\Sfr)$ by
    the prefix words of each $\Rfr_i$. This, together with the
    definition~\eqref{equ:rewrite_relation_induced} of the rewrite
    relation $\RewContext$ induced by $\Rew$ and the hypothesis of the
    statement of the lemma, implies that if $\Tfr$ and $\Tfr'$ are two
    binary trees such that $\Tfr \RewContext \Tfr'$,
    $\PrefixWord(\Tfr) > \PrefixWord\left(\Tfr'\right)$. This leads to
    the fact that any chain
    \begin{math}
        \Tfr_0 \RewContext \Tfr_1 \RewContext \Tfr_2 \RewContext \cdots
    \end{math}
    is finite since
    \begin{math}
        \Tfr_0 > \Tfr_1 > \Tfr_2 > \cdots
    \end{math}
    and there is a finite number of binary trees of a fixed arity.
    Therefore, $\RewContext$ is terminating.
\end{proof}
\medbreak

@@

Let us now go back on operads. Let $\Congr$ be an operad congruence of
$\Mag$. If $\Tfr$ is a binary tree, we denote by $[\Tfr]_{\Congr}$ the
$\Congr$-equivalence class of $\Tfr$. By definition, $[\Tfr]_{\Congr}$
is an element of the quotient operad
\begin{equation} \label{equ:quotient_operad_mag}
    \Oca := \Mag/_{\Congr}.
\end{equation}
A set of rewrite rules $\Rew$ on $\Mag$ is an \Def{orientation} of
$\Congr$ if $\RewContextRST$ and $\Congr$ are equal as binary relations,
where $\RewContext$ is the rewrite relation induced by $\Rew$. Moreover,
$\Rew$ is a \Def{convergent} (resp. \Def{terminating}, \Def{confluent})
orientation of $\Congr$ if $\RewContext$ is convergent (resp.
terminating, confluent). In this text, we call \Def{presentation} of a
quotient operad $\Oca$ of the form~\eqref{equ:quotient_operad_mag} the
data of a generating set for the operad congruence $\Congr$. Observe
that any orientation of $\Congr$ is a presentation of $\Oca$, so that
the above nomenclature (\Def{convergent}, \Def{terminating}, and
\Def{confluent}) still holds for presentations. A presentation is said
to be \Def{finite} if it is a finite set.
\medbreak
